---
title: "Apple Mobility Code - Oct 2nd"
author: "Tom Bird and Cerren Richards"
date: "02/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Set up the Apple data

```{r}
library(dplyr)
library(reshape2)
library(countrycode)
library(RCurl)
library(mgcv)
library(lubridate)
library(ggplot2)

apple = read.csv("applemobilitytrends-2020-08-26.csv") %>% 
  subset(grepl("country", geo_type) & transportation_type=="driving") %>%
  melt(id.vars=1:6) %>%
  mutate(variable=gsub("X", "", variable), country=region) %>%
  mutate(variable=gsub("\\.", "/", variable)) %>%
  mutate(Date=as.Date(variable, format="%Y/%m/%d")) %>%
  mutate(ISO3=countrycode(sourcevar=region, origin="country.name", destination="iso3c"))


```

### COVID-19 Confinement Data

The Oxford Covid-19 Government Response Tracker (OxCGRT) has compiled dates for eight different confinement types and their severity (0-4).

Data info: https://github.com/OxCGRT/covid-policy-tracker
```{r}

confinement <- getURL("https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest_withnotes.csv")


# Set up the confinement data for further analyses
conf <- read.csv(text=confinement) %>%
  mutate(julian=(strptime(Date, "%Y%m%d")$yday)+1) %>% # assign Julian Date
  mutate(Date=as.Date(strptime(Date, "%Y%m%d"))) %>% # Set as a date
  mutate(weekday=weekdays(Date)) %>% # Identify weekdays
  mutate(weekend=ifelse(grepl("Sat|Sun", weekday), 1,0)) %>% # Assign columns for weekends
  select(ISO3=CountryCode, RegionName, Date, julian, weekday, weekend, # Select columns
         C1_School.closing, C2_Workplace.closing, C3_Cancel.public.events,
         C4_Restrictions.on.gatherings, C5_Close.public.transport,
         C6_Stay.at.home.requirements, C7_Restrictions.on.internal.movement)

library(naniar)
# Remove subregions
conf<- conf %>% replace_with_na(replace = list(RegionName = ""))
conf<- conf[is.na(conf$RegionName),]

applemobconf=left_join(apple, conf)

# Sensitivity analysis to identify the best cut off dates
# We select May because this is when most countries begin to ease the restrictions
maystart <- 122 # Julian date for May 1st
maymid <- 136 # May 15
mayend <- 152 # May 30


# Subset the apple data
applestart <- subset(applemobconf,  julian <= maystart)
applemid <-   subset(applemobconf,  julian <= maymid)
appleend <-   subset(applemobconf,  julian <= mayend)

```

### Extract Model Values
```{r}
conftypes=c("C1_School.closing", "C2_Workplace.closing", "C3_Cancel.public.events",
            "C4_Restrictions.on.gatherings", "C5_Close.public.transport",
            "C6_Stay.at.home.requirements", "C7_Restrictions.on.internal.movement")


###_________________________________________
### MID MAY
###_________________________________________

country=ISO3=restriction=level=date=character()
aics=R2=t.lockdown=numeric()

country.results=subdat.list=chosen_mid=data.frame()
mod.list=country.results=list()

## Extract the model values for each country
for(i in unique(applemid$region)){
  for(j in conftypes){
    subdat=subset(applemid, region==i)  
    locklevels=unique(subdat[[j]])
    locklevels=locklevels[which(locklevels %in% c(1:10))]
    for(k in locklevels){
      subdat$lockdown=(subdat[[j]]==k)*1
      subdat$lockdown.type=j
      subdat$level=k
      mod=try(gam(value ~ lockdown + weekend + s(julian), data=subdat))
      if(!class(mod)[1]=="try-error"){
        country=c(country,i)
        ISO3=c(ISO3, subdat$ISO3[1])
        restriction=c(restriction,j)
        level=c(level,k)
        aics=c(aics,AIC(mod))
        R2=c(R2, summary(mod)$r.sq)
        t.lockdown=c(t.lockdown, summary(mod)$p.table["lockdown","t value"])
        date=c(date, subdat$variable[min(which(subdat$lockdown==1))])
        
        # save models and subdata
        mod.list[[paste(i,j,k, sep="_")]]=mod
        subdat.list=rbind(subdat.list, subset(subdat, select=c("country", "Date", "julian", "weekday", "weekend", "lockdown.type","level",  "value", "lockdown")))
        
      }
    }
  }
  
  # select country top model
  country.out=data.frame(country, ISO3, restriction,level, aics, R2, t.lockdown, date) %>%
    mutate(Date=ymd(date)) %>%
    mutate(model=paste(country, restriction, level, sep="_")) %>%
    subset(country==i) %>% 
    arrange(aics)
  
  country.results=rbind(country.results,country.out)
  
  country.out=subset(country.out, t.lockdown==min(t.lockdown))
  country.out$model = with(country.out, paste(country, restriction, level, sep="_"))
  
  # add top model outputs to list
  chosen_mid=rbind(chosen_mid,country.out)
  
  print(paste(i))
}


# Assign Julian date for the confinement date
chosen_mid$julian_lockdown <- (strptime(chosen_mid$Date, "%Y-%m-%d")$yday) + 1 

chosen_mid<- distinct(chosen_mid, country, .keep_all = TRUE)

# Find median AIC score for all countries
median(chosen_mid$aics) #864



###_________________________________________
### END MAY
###_________________________________________

country=ISO3=restriction=level=date=character()
aics=R2=t.lockdown=numeric()

country.results=subdat.list=chosen_end=data.frame()
mod.list=country.results=list()

## Extract the model values for each country
for(i in unique(appleend$region)){
  for(j in conftypes){
    subdat=subset(appleend, region==i)  
    locklevels=unique(subdat[[j]])
    locklevels=locklevels[which(locklevels %in% c(1:10))]
    for(k in locklevels){
      subdat$lockdown=(subdat[[j]]==k)*1
      subdat$lockdown.type=j
      subdat$level=k
      mod=try(gam(value ~ lockdown + weekend + s(julian), data=subdat))
      if(!class(mod)[1]=="try-error"){
        country=c(country,i)
        ISO3=c(ISO3, subdat$ISO3[1])
        restriction=c(restriction,j)
        level=c(level,k)
        aics=c(aics,AIC(mod))
        R2=c(R2, summary(mod)$r.sq)
        t.lockdown=c(t.lockdown, summary(mod)$p.table["lockdown","t value"])
        date=c(date, subdat$variable[min(which(subdat$lockdown==1))])
        
        # save models and subdata
        mod.list[[paste(i,j,k, sep="_")]]=mod
        subdat.list=rbind(subdat.list, subset(subdat, select=c("country", "Date", "julian", "weekday", "weekend", "lockdown.type","level",  "value", "lockdown")))
        
      }
    }
  }
  
  # select country top model
  country.out=data.frame(country, ISO3, restriction,level, aics, R2, t.lockdown, date) %>%
    mutate(Date=ymd(date)) %>%
    mutate(model=paste(country, restriction, level, sep="_")) %>%
    subset(country==i) %>% 
    arrange(aics)
  
  country.results=rbind(country.results,country.out)
  
  country.out=subset(country.out, t.lockdown==min(t.lockdown))
  country.out$model = with(country.out, paste(country, restriction, level, sep="_"))
  
  # add top model outputs to list
  chosen_end=rbind(chosen_end,country.out)
  
  print(paste(i))
}


# Assign Julian date for the confinement date
chosen_end$julian_lockdown <- (strptime(chosen_end$Date, "%Y-%m-%d")$yday) + 1 

chosen_end<- distinct(chosen_end, country, .keep_all = TRUE)

# Find median AIC score for all countries
median(chosen_end$aics) #972


###_________________________________________
### START MAY
###_________________________________________

country=ISO3=restriction=level=date=character()
aics=R2=t.lockdown=numeric()

country.results=subdat.list=chosen_start=data.frame()
mod.list=country.results=list()

## Extract the model values for each country
for(i in unique(applestart$region)){
  for(j in conftypes){
    subdat=subset(applestart, region==i)  
    locklevels=unique(subdat[[j]])
    locklevels=locklevels[which(locklevels %in% c(1:10))]
    for(k in locklevels){
      subdat$lockdown=(subdat[[j]]==k)*1
      subdat$lockdown.type=j
      subdat$level=k
      mod=try(gam(value ~ lockdown + weekend + s(julian), data=subdat))
      if(!class(mod)[1]=="try-error"){
        country=c(country,i)
        ISO3=c(ISO3, subdat$ISO3[1])
        restriction=c(restriction,j)
        level=c(level,k)
        aics=c(aics,AIC(mod))
        R2=c(R2, summary(mod)$r.sq)
        t.lockdown=c(t.lockdown, summary(mod)$p.table["lockdown","t value"])
        date=c(date, subdat$variable[min(which(subdat$lockdown==1))])
        
        # save models and subdata
        mod.list[[paste(i,j,k, sep="_")]]=mod
        subdat.list=rbind(subdat.list, subset(subdat, select=c("country", "Date", "julian", "weekday", "weekend", "lockdown.type","level",  "value", "lockdown")))
        
      }
    }
  }
  
  # select country top model
  country.out=data.frame(country, ISO3, restriction,level, aics, R2, t.lockdown, date) %>%
    mutate(Date=ymd(date)) %>%
    mutate(model=paste(country, restriction, level, sep="_")) %>%
    subset(country==i) %>% 
    arrange(aics)
  
  country.results=rbind(country.results,country.out)
  
  country.out=subset(country.out, t.lockdown==min(t.lockdown))
  country.out$model = with(country.out, paste(country, restriction, level, sep="_"))
  
  # add top model outputs to list
  chosen_start=rbind(chosen_start,country.out)
  
  print(paste(i))
}


# Assign Julian date for the confinement date
chosen_start$julian_lockdown <- (strptime(chosen_start$Date, "%Y-%m-%d")$yday) + 1 

chosen_start<- distinct(chosen_start, country, .keep_all = TRUE)

# Find median AIC score for all countries
median(chosen_start$aics) #783


```


### Plot to visualise the dates selected
```{r}

# Join the lockdown dates to the full Google data for plotting
plots <- left_join(applestart, chosen_start, by = c("country", "ISO3"))

library(ggplot2)
library(ggforce)
## PLOT
ggplot(data = plots, aes(x=julian, y=value)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 1)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")

ggplot(data = plots, aes(x=julian, y=value)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 2)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")


```

### Find the date for countries that dont work

The models above did not select appropriate dates. We therefore manually select the dates for the following countries:

Cambodia, Italy, Japan, Slovakia, Turkey

```{r}

# Subset the countries that the model did not select an appropriate confinement date

# Set Julian date
country.results$julian_lockdown <- (strptime(country.results$Date, "%Y-%m-%d")$yday) + 1 

## Cambodia_C4_Restrictions.on.gatherings_4 = Julian day = 77
ggplot(data = filter(subdat.list, country == "Cambodia"), aes(x=julian, y=value)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Cambodia"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Cambodia"))


## Italy_C5_Close.public.transport_1 = Julian day = 71
ggplot(data = filter(subdat.list, country == "Italy"), aes(x=julian, y=value)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Italy"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Italy"))


## Japan_C6_Stay.at.home.requirements_1 = Julian day = 98
ggplot(data = filter(subdat.list, country == "Japan"), aes(x=julian, y=value)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Japan"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Japan"))


## Slovakia_C3_Cancel.public.events_2 = Julian day = 70
ggplot(data = filter(subdat.list, country == "Slovakia"), aes(x=julian, y=value)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Slovakia"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Slovakia"))

## Turkey_C7_Restrictions.on.internal.movement_1 = Julian day = 78
ggplot(data = filter(subdat.list, country == "Turkey"), aes(x=julian, y=value)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Turkey"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Turkey"))


corrections <- tibble(country = c("Cambodia","Italy", "Japan", "Slovakia","Turkey"), 
                      model = c("Cambodia_C4_Restrictions.on.gatherings_4",
                                "Italy_C5_Close.public.transport_1",
                                "Japan_C6_Stay.at.home.requirements_1",
                                "Slovakia_C3_Cancel.public.events_2",
                                "Turkey_C7_Restrictions.on.internal.movement_1"),
                      julian_lockdown = c(77,71,98,70,78),
                      aics = NA,
                      R2 = NA,
                      t.lockdown = NA)



# Add the ISO3 for the country names
corrections$ISO3 <- countrycode(corrections$country, origin = 'country.name', destination = 'iso3c')

# Loop through and extract the values from the selected models
for (i in 1:8) { 
  
  corrections$t.lockdown[i] <- summary(mod.list[[corrections$model[i]]])$p.table["lockdown","t value"]
  corrections$R2[i] <- summary(mod.list[[corrections$model[i]]])$r.sq
  corrections$aics[i] <- AIC(mod.list[[corrections$model[i]]])
}


```


### Calculate the mean percent change following introduction of confinement measures

```{r}
library(tidyr)
# Update the corrected model values
chosen_start <- bind_rows(corrections, chosen_start) %>%
  distinct(country, .keep_all = T)


apple_all <- left_join(applestart %>% dplyr::select(country, value, ISO3, julian, weekend),
                        chosen_start %>% dplyr::select(country, ISO3, model, julian_lockdown), 
                        by = c("country", "ISO3"))

# Remove Taiwan because the country does not show a response to lockdown
apple_all <- apple_all %>% filter(!country ==  "Taiwan")

# Assign before and after pandemic
apple_all <- mutate(apple_all, lockdown = ifelse(julian >= julian_lockdown, "After", "Before"))

# Set levels for the model
apple_all$lockdown <- factor(apple_all$lockdown, levels = c("Before","After"))

# Apple data show change above or below 100 rather than 0, so calculate the percent change from 100
apple_all$percent <- apple_all$value - 100

# get median change after pandemic 
mean_after<- apple_all %>% group_by(ISO3, lockdown) %>% 
  summarise(percent = mean(percent))

# Filter only after the pandemic declared
mean_after <- filter(mean_after, lockdown == "After")

chosen_start <- left_join(chosen_start, mean_after)

# save chosen models
write.csv(chosen_start, file="Oct2nd.Applemob.chosen.csv") 


```

### Modelling the global percent change

```{r}

mod.drive <- gamm(data = apple_all, percent ~ lockdown + weekend + s(julian), random = list(country=~1))

summary(mod.drive$gam)
## coefficient = -41.2731
## error = 0.8939
## tval = -46.173
## n = 6820

```

