---
title: "PAN-Environment Google mobility code"
authors: "Cerren Richards & Tom Bird"
date: "02/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we use the Google mobility data to find the t-value for all countries before and after COVID-19 confinements.

### Google Data

**Data Description**:

* How visits and length of stay at different places change compared to a baseline.
    - Parks: national parks, public beaches, marinas, dog parks, plazas, and public gardens.
    - Residential: places of residence
    
**Baseline**:

* Median value, for the corresponding day of the week, during the 5- week period Jan 3–Feb 6, 2020

**Download data and methods**:

* https://www.google.com/covid19/mobility/
* https://support.google.com/covid19-mobility/answer/9824897?hl=en&ref_topic=9822927

```{r, eval=FALSE}

# Read in google data
google <- read.csv("Global_Mobility_Report.csv")

# R reads Namibia's iso2 code is as an "is.na" object, so we will rename it to NA
google$country_region_code[is.na(google$country_region_code)] <-"NA"

# Add the country codes so we ensure everything matches 
library(countrycode)
google$ISO2 <- google$country_region_code
google$ISO3 <- countrycode(google$ISO2, origin = 'iso2c', destination = 'iso3c')
google$country <- countrycode(google$ISO2, origin = 'iso2c', destination = 'country.name')

# Add Julian date
google$julian<- (strptime(google$date, "%Y-%m-%d")$yday) + 1 # need to add 1 day because 2020 is a leap year

library(naniar)
# remove the subregions and only keep the overall country pattern
google<- google %>% replace_with_na(replace = list(sub_region_1 = ""))
google<- google[is.na(google$sub_region_1),]

library(dplyr)
# Select the columns that will be used in further analyses
google <- google %>% select(country, ISO3, date, julian, 
                            residential = residential_percent_change_from_baseline, 
                            parks = parks_percent_change_from_baseline)

# Sensitivity analysis to identify the best cut off dates
# We select May because this is when most countries begin to ease the restrictions
maystart <- 122 # Julian date for May 1st
maymid <- 136 # May 15
mayend <- 152 # May 30

# Subset the google data
googlestart <- subset(google,  julian <= maystart)
googlemid <- subset(google,  julian <= maymid)
googleend <- subset(google,  julian <= mayend)


```


### COVID-19 Confinement Data

The Oxford Covid-19 Government Response Tracker (OxCGRT) has compiled dates for eight different confinement types and their severity (0-4).

Data info: https://github.com/OxCGRT/covid-policy-tracker

```{r, eval=FALSE}
library(RCurl)

# Call the OxCGRT data from GitHub
confinement <- getURL("https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest_withnotes.csv")

# Set up the confinement data for further analyses
conf <- read.csv(text=confinement) %>% 
  mutate(julian=(strptime(Date, "%Y%m%d")$yday)+1) %>% # assign Julian Date
  mutate(Date=as.Date(strptime(Date, "%Y%m%d"))) %>% # Set as a date 
  mutate(weekday=weekdays(Date)) %>% # Identify weekdays
  mutate(weekend=ifelse(grepl("Sat|Sun", weekday), 1,0)) %>% # Assign columns for weekends
  select(ISO3=CountryCode, RegionName, Date, julian, weekday, weekend, # Select columns
         C1_School.closing, C2_Workplace.closing, C3_Cancel.public.events,
         C4_Restrictions.on.gatherings, C5_Close.public.transport,
         C6_Stay.at.home.requirements, C7_Restrictions.on.internal.movement)

# Remove subregions
conf<- conf %>% replace_with_na(replace = list(RegionName = ""))
conf<- conf[is.na(conf$RegionName),]


# Join the Google data with the confinement data
googlestart <- left_join(googlestart, conf, by = c("ISO3", "julian"))
googlemid <- left_join(googlemid, conf, by = c("ISO3", "julian"))
googleend <- left_join(googleend, conf, by = c("ISO3", "julian"))


```


### SENSITIVITY ANALYSIS and Extract t-statistic for each country

We divided each country’s mobility data into ‘before’ and ‘after’ periods for each confinement type and level, ran generalised additive models to determine which division best characterised the observed changes in driving patterns. 

```{r, eval = FALSE}

# List seven of the confinement types 
conftypes=c("C1_School.closing", "C2_Workplace.closing", "C3_Cancel.public.events",
            "C4_Restrictions.on.gatherings", "C5_Close.public.transport",
            "C6_Stay.at.home.requirements", "C7_Restrictions.on.internal.movement")

###_________________________________________
### MID MAY
###_________________________________________

# Create empty spaces for results to be pasted
country=ISO3=restriction=level=date=character() # set as character
aics=R2=t.lockdown=pval=numeric() # set as numeric

# Create empty dataframes and lists for model outputs
country.results=subdat.list=chosen_mid=data.frame()
mod.list=country.results=list()

library(mgcv)
library(lubridate)

# Run loop to extract the t-statistic
for(i in unique(googlemid$country)){
  for(j in conftypes){
    subdat=subset(googlemid, country==i)  
    locklevels=unique(subdat[[j]])
    locklevels=locklevels[which(locklevels %in% c(1:10))]
    for(k in locklevels){
      subdat$lockdown=(subdat[[j]]==k)*1
      subdat$lockdown.type=j
      subdat$level=k
      mod=try(gam(residential ~ lockdown + weekend + s(julian), data=subdat))
      if(!class(mod)[1]=="try-error"){
        country=c(country,i)
        ISO3=c(ISO3, subdat$ISO3[1])
        restriction=c(restriction,j)
        level=c(level,k)
        aics=c(aics,AIC(mod))
        R2=c(R2, summary(mod)$r.sq)
        t.lockdown=c(t.lockdown, summary(mod)$p.table["lockdown","t value"])
        pval=c(pval, summary(mod)$p.table["lockdown","Pr(>|t|)"])
        date=c(date, subdat$date[min(which(subdat$lockdown==1))])
      

        # save models and subdata
        mod.list[[paste(i,j,k, sep="_")]]=mod
        subdat.list=rbind(subdat.list, subset(subdat, 
                                              select=c("country", "date", "julian", "weekday",
                                                       "lockdown.type","level",  "residential",
                                                        "lockdown")))
        
      }
    }
  }
  
  # select country top model
  country.out=data.frame(country, ISO3, restriction, level, aics, R2, t.lockdown, pval, date) %>%
    mutate(Date=ymd(date)) %>%
    subset(country==i) %>% 
    arrange(aics)
  
  country.results=rbind(country.results,country.out)
  
  country.out=subset(country.out, t.lockdown==max(t.lockdown))
  country.out$model = with(country.out, paste(country, restriction, level, sep="_"))
  
  # add top model outputs to list
  chosen_mid=rbind(chosen_mid,country.out)
  
  print(paste(i))
}

# Assign Julian date for the confinement date
chosen_mid$julian_lockdown <- (strptime(chosen_mid$date, "%Y-%m-%d")$yday) + 1 

chosen_mid<- distinct(chosen_mid, country, .keep_all = TRUE)

# Find median AIC score for all countries
median(chosen_mid$aics) #447

###_________________________________________
### END MAY
###_________________________________________


# Create empty spaces for results to be pasted
country=ISO3=restriction=level=date=character() # set as character
aics=R2=t.lockdown=pval=numeric() # set as numeric

# Create empty dataframes and lists for model outputs
country.results=subdat.list=chosen_end=data.frame()
mod.list=country.results=list()


# Run loop to extract the t-statistic
for(i in unique(googleend$country)){
  for(j in conftypes){
    subdat=subset(googleend, country==i)  
    locklevels=unique(subdat[[j]])
    locklevels=locklevels[which(locklevels %in% c(1:10))]
    for(k in locklevels){
      subdat$lockdown=(subdat[[j]]==k)*1
      subdat$lockdown.type=j
      subdat$level=k
      mod=try(gam(residential ~ lockdown + weekend + s(julian), data=subdat))
      if(!class(mod)[1]=="try-error"){
        country=c(country,i)
        ISO3=c(ISO3, subdat$ISO3[1])
        restriction=c(restriction,j)
        level=c(level,k)
        aics=c(aics,AIC(mod))
        R2=c(R2, summary(mod)$r.sq)
        t.lockdown=c(t.lockdown, summary(mod)$p.table["lockdown","t value"])
        pval=c(pval, summary(mod)$p.table["lockdown","Pr(>|t|)"])
        date=c(date, subdat$date[min(which(subdat$lockdown==1))])
      

        # save models and subdata
        mod.list[[paste(i,j,k, sep="_")]]=mod
        subdat.list=rbind(subdat.list, subset(subdat, 
                                              select=c("country", "date", "julian", "weekday",
                                                       "lockdown.type","level",  "residential",
                                                        "lockdown")))
        
      }
    }
  }
  
  # select country top model
  country.out=data.frame(country, ISO3, restriction, level, aics, R2, t.lockdown, pval, date) %>%
    mutate(Date=ymd(date)) %>%
    subset(country==i) %>% 
    arrange(aics)
  
  country.results=rbind(country.results,country.out)
  
  country.out=subset(country.out, t.lockdown==max(t.lockdown))
  country.out$model = with(country.out, paste(country, restriction, level, sep="_"))
  
  # add top model outputs to list
  chosen_end=rbind(chosen_end,country.out)
  
  print(paste(i))
}

# Assign Julian date for the confinement date
chosen_end$julian_lockdown <- (strptime(chosen_end$date, "%Y-%m-%d")$yday) + 1 


chosen_end<- distinct(chosen_end, country, .keep_all = TRUE)

# Find median AIC score for all countries
median(chosen_end$aics) #525


###_________________________________________
### START MAY
###_________________________________________

# Create empty spaces for results to be pasted
country=ISO3=restriction=level=date=character() # set as character
aics=R2=t.lockdown=pval=numeric() # set as numeric

# Create empty dataframes and lists for model outputs
country.results=subdat.list=chosen_start=data.frame()
mod.list=country.results=list()

# Run loop to extract the t-statistic
for(i in unique(googlestart$country)){
  for(j in conftypes){
    subdat=subset(googlestart, country==i)  
    locklevels=unique(subdat[[j]])
    locklevels=locklevels[which(locklevels %in% c(1:10))]
    for(k in locklevels){
      subdat$lockdown=(subdat[[j]]==k)*1
      subdat$lockdown.type=j
      subdat$level=k
      mod=try(gam(residential ~ lockdown + weekend + s(julian), data=subdat))
      if(!class(mod)[1]=="try-error"){
        country=c(country,i)
        ISO3=c(ISO3, subdat$ISO3[1])
        restriction=c(restriction,j)
        level=c(level,k)
        aics=c(aics,AIC(mod))
        R2=c(R2, summary(mod)$r.sq)
        t.lockdown=c(t.lockdown, summary(mod)$p.table["lockdown","t value"])
        pval=c(pval, summary(mod)$p.table["lockdown","Pr(>|t|)"])
        date=c(date, subdat$date[min(which(subdat$lockdown==1))])
      

        # save models and subdata
        mod.list[[paste(i,j,k, sep="_")]]=mod
        subdat.list=rbind(subdat.list, subset(subdat, 
                                              select=c("country", "date", "julian", "weekday",
                                                       "lockdown.type","level",  "residential",
                                                        "lockdown")))
        
      }
    }
  }
  
  # select country top model
  country.out=data.frame(country, ISO3, restriction, level, aics, R2, t.lockdown, pval, date) %>%
    mutate(date=ymd(date)) %>%
    subset(country==i) %>% 
    arrange(aics)
  
  country.results=rbind(country.results,country.out)
  
  country.out=subset(country.out, t.lockdown==max(t.lockdown))
  country.out$model = with(country.out, paste(country, restriction, level, sep="_"))
  
  # add top model outputs to list
  chosen_start=rbind(chosen_start,country.out)
  
  print(paste(i))
}

# Assign Julian date for the confinement date
chosen_start$julian_lockdown <- (strptime(chosen_start$date, "%Y-%m-%d")$yday) + 1 

# Remove duplicated countries (that have the same AIC score for multiple confinement types)
chosen_start<- distinct(chosen_start, country, .keep_all = TRUE)

# Find median AIC score for all countries
median(chosen_start$aics) #379


```


### Plot the individual country trends with lockdown date

```{r, eval = FALSE}

# Join the lockdown dates to the full Google data for plotting
googleplots <- left_join(googlestart, chosen_start, by = c("country", "ISO3"))

library(ggplot2)
library(ggforce)
## PLOT
ggplot(data = googleplots, aes(x=julian, y=residential)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 1)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")

ggplot(data = googleplots, aes(x=julian, y=residential)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 2)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")

ggplot(data = googleplots, aes(x=julian, y=residential)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 3)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")


```


### Find the date for countries that dont work

The models above did not select appropriate dates. We therefore manually select the dates for the following countries:

"Bahrain", "Belize", "Cambodia","Hong Kong SAR China", "Italy",
"Indonesia", "Mongolia", "Portugal","South Korea", "Taiwan", "Vietnam","Yemen"


```{r}
# Subset the countries that the model did not select an appropriate confinement date

# Set Julian date
country.results$julian_lockdown <- (strptime(country.results$date, "%Y-%m-%d")$yday) + 1 

## Bahrain_C2_Workplace.closing_2 = Julian day = 78
ggplot(data = filter(subdat.list, country == "Bahrain"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Bahrain"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Bahrain"))


## Belize_C1_School.closing_3 = Julian day = 80
ggplot(data = filter(subdat.list, country == "Belize"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Belize"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Belize"))


## Cambodia_C2_Workplace.closing_2 = Julian day = 77
ggplot(data = filter(subdat.list, country == "Cambodia"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Cambodia"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Cambodia"))


## Hong Kong SAR China = NOT SUITABLE 
ggplot(data = filter(subdat.list, country == "Hong Kong SAR China"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Hong Kong SAR China"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Hong Kong SAR China"))


## Italy_C5_Close.public.transport_1 = Julian day = 71
ggplot(data = filter(subdat.list, country == "Italy"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Italy"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Italy"))


## Indonesia_C5_Close.public.transport_1 = Julian day = 76
ggplot(data = filter(subdat.list, country == "Indonesia"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Indonesia"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Indonesia"))


## Mongolia_C7_Restrictions.on.internal.movement_2 = Julian day = 52  not sure if can use
ggplot(data = filter(subdat.list, country == "Mongolia"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Mongolia"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Mongolia"))


## Portugal_C2_Workplace.closing_2 = Julian day = 72
ggplot(data = filter(subdat.list, country == "Portugal"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Portugal"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Portugal"))


## South Korea = cant use?
ggplot(data = filter(subdat.list, country == "South Korea"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "South Korea"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "South Korea"))

## Taiwan = cant use
ggplot(data = filter(subdat.list, country == "Taiwan"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Taiwan"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Taiwan"))


## Vietnam_C2_Workplace.closing_3 = Julian day = 87
ggplot(data = filter(subdat.list, country == "Vietnam"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Vietnam"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Vietnam"))


## Yemen_C3_Cancel.public.events_2 = Julian day = 77 - cant use??
ggplot(data = filter(subdat.list, country == "Yemen"), aes(x=julian, y=residential)) + 
  geom_line() + 
  geom_vline(aes(xintercept=julian_lockdown, colour=restriction), 
             data = filter(country.results, country == "Yemen"))+
  geom_text(aes(x=julian_lockdown, y=10 + 2*as.numeric(level), label=level), 
            data = filter(country.results, country == "Yemen"))



# We exclude "Hong Kong SAR China","Mongolia", "South Korea", "Taiwan", "Yemen"
# because they do not show an obvious change in lockdown date

# Extract the model values for the countries that need corrected confinement dates
corrections <- tibble(country = c("Bahrain", "Belize", "Cambodia", 
                                  "Italy", "Indonesia", "Portugal",
                                  "Vietnam"), 
                      model = c("Bahrain_C2_Workplace.closing_2", "Belize_C1_School.closing_3",
                                "Cambodia_C2_Workplace.closing_2", "Italy_C5_Close.public.transport_1",
                                "Indonesia_C5_Close.public.transport_1", "Portugal_C2_Workplace.closing_2",
                                "Vietnam_C2_Workplace.closing_3"),
                      julian_lockdown = c(78, 80, 77, 71, 76, 72, 87),
                      aics = NA,
                      R2 = NA,
                      t.lockdown = NA)

# Add the ISO3 for the country names
corrections$ISO3 <- countrycode(corrections$country, origin = 'country.name', destination = 'iso3c')

# Loop through and extract the values from the selected models
for (i in 1:7) { 

corrections$t.lockdown[i] <- summary(mod.list[[corrections$model[i]]])$p.table["lockdown","t value"]
corrections$pval[i] <- summary(mod.list[[corrections$model[i]]])$p.table["lockdown","Pr(>|t|)"]
corrections$R2[i] <- summary(mod.list[[corrections$model[i]]])$r.sq
corrections$aics[i] <- AIC(mod.list[[corrections$model[i]]])
}

library(tidyr)
# Update the corrected model values
chosen_residential <- bind_rows(corrections, chosen_start)
chosen_residential <- distinct(chosen_residential, country, .keep_all = T)

# Remove countries
chosen_residential <- chosen_residential %>% filter(!country %in% c("Hong Kong SAR China","Mongolia", "South Korea", "Taiwan", "Yemen"))

```



### Extract t-statistic PARKS

```{r}

# Join the confinement dates to the google data
googlepark <- left_join(googlestart %>% dplyr::select(country:weekend),
                 chosen_residential %>% dplyr::select(country, ISO3, model, julian_lockdown), 
                 by = c("country", "ISO3"))

# Remove countries
googlepark <- googlepark %>% filter(!country %in% c("Hong Kong SAR China","Mongolia", "South Korea", "Taiwan", "Yemen"))


# Assign whether the date is before or after lockdown
googlepark <- googlepark %>% group_by(country) %>% 
  mutate(lockdown = ifelse(julian >= julian_lockdown, "After", "Before"))

# Set levels for the model
googlepark$lockdown <- factor(googlepark$lockdown, levels = c("Before","After"))

# Make a list of all the countries with no subregions
# This will be called on in the loop below and the tvalues will be placed here
chosen_parks <- chosen_residential %>% select(country, ISO3, model, date, julian_lockdown)


# PARKS MODEL

#Assign NA to all countries
chosen_parks$t.lockdown <- NA
chosen_parks$pval <- NA
chosen_parks$R2 <- NA
chosen_parks$aics <- NA

# Run GAM model for each country in the no_sub_countries dataframe
for (i in 1:nrow(chosen_parks)) { 
 tryCatch({

model <- gam(data=filter(googlepark, country == chosen_parks$country[i]), 
              parks ~ lockdown + weekend + s(julian))

# Extract the t-value and paste it into the no_sub_countries dataframe
chosen_parks$t.lockdown[i] <- summary(model)$p.table["lockdownAfter","t value"]
chosen_parks$pval[i] <- summary(model)$p.table["lockdownAfter","Pr(>|t|)"]
chosen_parks$R2[i] <- summary(model)$r.sq
chosen_parks$aics[i] <- AIC(model)

  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

```



### PARKS Plot the individual country trends with lockdown date

```{r, eval = FALSE}

# Join the lockdown dates to the full Google data for plotting
#googleplots <- left_join(google, chosen, by = "country")

library(ggforce)
## PLOT
ggplot(data = googlepark, aes(x=julian, y=parks)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 1)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")

ggplot(data = googleplots, aes(x=julian, y=parks)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 2)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")

ggplot(data = googlepark, aes(x=julian, y=parks)) + 
  geom_line() + 
  facet_wrap_paginate(~country, ncol = 7, nrow = 7, scale="free_y", page = 3)+
  geom_vline(aes(xintercept = julian_lockdown), colour = "red")


```



### Calculate the mean percent change following introduction of confinement measures

```{r, eval=FALSE}

# Assign before and after pandemic
#googleplots <- mutate(googleplots, lockdown = ifelse(julian >= julian_lockdown, "After", "Before"))

# get median change after pandemic 
mean_after<- googlepark %>% group_by(ISO3, lockdown) %>% 
  summarise(mean_res = mean(residential),
            mean_park = mean(parks))

# Filter only after the pandemic declared
mean_after <- filter(mean_after, lockdown == "After")

# Join the percent change
chosen_parks <- left_join(chosen_parks, 
                 mean_after %>% dplyr::select(percent = mean_park), 
                 by = c("ISO3"))

chosen_residential <- left_join(chosen_residential, 
                 mean_after %>% dplyr::select(percent = mean_res), 
                 by = c("ISO3"))


# Revere the direction of residential 
chosen_residential$percent <- chosen_residential$percent*-1 
chosen_residential$t.lockdown <- chosen_residential$t.lockdown*-1 

```


### Plot Percentage change
```{r}

# Define increase or decrease
median_after <- mutate(median_after, increase_res = ifelse(med_res > 0 , "Increase", "Decrease"))
median_after <- mutate(median_after, increase_park = ifelse(med_park > 0 , "Increase", "Decrease"))

#Join to `world` dataframe (it's in the above R chunk)
median_after <- left_join(world, median_after)


# TWO COLOURS
ggplot() +
  theme_void(base_size = 18)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   geom_sf(data = median_after, aes(fill = increase_park),  colour = "grey20", size = 0.25)+
  scale_y_continuous(expand = c(0,0), limits = c(-60, 85))+
  scale_x_continuous(expand = c(0,0), limits = c(-180, 180))+
  labs(fill = "change at parks")+
 theme(legend.position = "bottom", legend.box = "vertical")

#CONTINUOUS COLOURS
ggplot() +
  theme_void(base_size = 15)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   geom_sf(data = median_after, aes(fill = med_res),  colour = "grey20", size = 0.25)+
  scale_y_continuous(expand = c(0,0), limits = c(-60, 85))+
  scale_x_continuous(expand = c(0,0), limits = c(-180, 180))+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())+
  theme(legend.position = "bottom", legend.box = "vertical")+
  scale_fill_viridis(option = "D", na.value="grey90",direction = 1,
  guide = guide_colorbar (title.position = 'top', title.hjust = 0.5, label.hjust = 0.5,
  barheight = unit(3, units = "mm"), barwidth = unit(100, units = "mm"))) +
  labs(fill = "Percent change - residential") 


ggplot() +
  theme_void(base_size = 15)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   geom_sf(data = median_after, aes(fill = med_park),  colour = "grey20", size = 0.25)+
  scale_y_continuous(expand = c(0,0), limits = c(-60, 85))+
  scale_x_continuous(expand = c(0,0), limits = c(-180, 180))+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())+
  theme(legend.position = "bottom", legend.box = "vertical")+
  scale_fill_viridis(option = "D", na.value="grey90", direction = 1,
  guide = guide_colorbar (title.position = 'top', title.hjust = 0.5, label.hjust = 0.5,
  barheight = unit(3, units = "mm"), barwidth = unit(100, units = "mm"))) +
  labs(fill = "Percent change - parks") 

#Percent increase at places of residence after confinement




```


## Sort the data ready for the stacked plot
```{r}
# assign names to the data
chosen_parks$type <- "Parks & Beaches"
chosen_residential$type <- "Residential"

# Join the two google data (parks and residential)
stacked <- bind_rows(chosen_residential, chosen_parks)
stacked <- stacked %>% select(country, ISO3, t.lockdown, percent, type)


# Read marine traffic data
boat <- read.csv("eez_ais_change.csv")

# Filter out the territories
boat <- boat %>% filter(TERRITORY1 == SOVEREIGN1)

# Select the data to keep
boat <- boat %>% select(country = TERRITORY1, 
                        percent = change, 
                        t.lockdown = t_after, 
                        ISO3 = ISO_SOV1) %>%
    distinct(country, .keep_all = T)

# Assign name
boat$type <- "Marine Traffic"

# create a new dataframe
stacked <- bind_rows(stacked, boat)


## Apple data
apple <- read.csv("Oct2nd.Applemob.chosen.csv")
apple$type <- "Driving"

# Select data
apple <- apple %>% select(country, percent,t.lockdown,ISO3, type)
stacked <- bind_rows(stacked, apple)


## Air Traffic
air <- read.csv("Airtraffic.effects.csv")
air$type <- "Air Traffic"
air$country <- countrycode(air$ISO3, origin = 'iso3c', destination = 'country.name')
air <- air %>% select(country, ISO3, percent = total.Perc.chg, type)
air$percent <- air$percent*100
air <- air %>% filter(country != "Egypt")
stacked <- bind_rows(stacked, air)


library(ggplot2);library(viridis); library(ggpubr);library(sf);library("rnaturalearth"); library("rnaturalearthdata")
# extract continent info
# data for the world
world <- ne_countries(scale = "medium", returnclass = "sf")
world <- world %>% rename(ISO3 = iso_a3) 
region_un <- world$region_un
ISO3 <- world$ISO3
continent<- tibble(ISO3, region_un)

# join to stacked data
stacked <- left_join(stacked, continent, by = "ISO3") 

# assign Tuvalu to Oceania
stacked$region_un[is.na(stacked$region_un)] <-"Oceania"


# Only keep the countries with all five mobility types
number <- stacked %>% count(country, sort = TRUE)
stacked <- left_join(stacked, number) 
stacked_all <- stacked
stacked <-stacked %>% filter(n == 5)


```

## Create stacked bar plot for the main manuscript

```{r}

# Sum the percentages to allow the ggplot code to plot from highest to lowest
stacked_sum <- stacked %>% group_by(region_un, country) %>% summarise(sum = sum(percent))
stacked <- left_join(stacked, stacked_sum)

# Reorder the mobility types
stacked$type <- factor(stacked$type, levels = c("Residential", "Parks & Beaches", "Driving",
                                              "Marine Traffic","Air Traffic"))

# continent + flat
ggplot(data = stacked, aes(x = reorder(country, -sum), y=percent, fill=type)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE, direction = -1)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
  facet_grid(.~region_un, scales = "free", space = "free")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
    theme(strip.text.x = element_blank(), 
          strip.background = element_blank())

# Save
ggsave("stacked.pdf", 
       dpi = 600, 
       width = 250, height = 120, unit = "mm")

```


## Plot for supplementary
Plot the percentage change after confinement for each country.

```{r}

# PARKS
chosen_parks <- left_join(chosen_parks, continent, by = "ISO3") 

ggplot(data = chosen_parks, aes(x = reorder(country, -percent), y=percent)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
  facet_grid(.~region_un, scales = "free", space = "free")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
  theme(strip.text.x = element_blank(), 
          strip.background = element_blank()) 

ggsave("Parks.pdf", 
       dpi = 600, 
       width = 400, height = 120, unit = "mm")



# RESIDENTIAL
chosen_residential <- left_join(chosen_residential, continent, by = "ISO3") 

ggplot(data = chosen_residential, aes(x = reorder(country, -percent), y=percent)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
  facet_grid(.~region_un, scales = "free", space = "free")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
  theme(strip.text.x = element_blank(), 
          strip.background = element_blank()) 

ggsave("Residential.pdf", 
       dpi = 600, 
       width = 600, height = 120, unit = "mm")


# AIR TRAFFIC
air <- left_join(air, continent, by = "ISO3") 

ggplot(data = air, aes(x = reorder(country, -percent), y=percent)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
  facet_grid(.~region_un, scales = "free", space = "free")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
  theme(strip.text.x = element_blank(), 
          strip.background = element_blank()) 

ggsave("Air Traffic.pdf", 
       dpi = 600, 
       width = 350, height = 120, unit = "mm")


# DRIVING
apple <- left_join(apple, continent, by = "ISO3") 

ggplot(data = apple, aes(x = reorder(country, -percent), y=percent)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
  facet_grid(.~region_un, scales = "free", space = "free")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
  theme(strip.text.x = element_blank(), 
          strip.background = element_blank()) 

ggsave("Driving.pdf", 
       dpi = 600, 
       width = 350, height = 120, unit = "mm")


# MARINE TRAFFIC

# Read marine traffic data
boat <- read.csv("eez_ais_change.csv")

# Filter out the territories
boat <- boat %>% filter(TERRITORY1 == SOVEREIGN1)

boat <- boat %>% select(country = TERRITORY1, 
                        percent = change, 
                        t.lockdown = t_after, 
                        ISO3 = ISO_SOV1) %>%
    distinct(country, .keep_all = T)

boat <- left_join(boat, continent, by = "ISO3") 
boat$region_un[is.na(boat$region_un)] <-"Oceania"


ggplot(data = boat, aes(x = reorder(country, -percent), y=percent)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
    facet_grid(.~region_un, scales = "free", space = "free")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
  theme(strip.text.x = element_blank(), 
          strip.background = element_blank()) 

ggsave("Marine Traffic.pdf", 
       dpi = 600, 
       width = 750, height = 160, unit = "mm")


## Marine Traffic Territories

# Read marine traffic data
boat <- read.csv("eez_ais_change.csv")

# Filter out the territories
boat <- boat %>% filter(TERRITORY1 != SOVEREIGN1)

boat <- boat %>% select(country = TERRITORY1, 
                        percent = change, 
                        t.lockdown = t_after, 
                        ISO3 = ISO_SOV1) %>%
    distinct(country, .keep_all = T)

boat <- boat %>% filter(country != "Heard and McDonald Islands")
    
boat <- left_join(boat, continent, by = "ISO3") 


    ggplot(data = boat, aes(x = reorder(country, -percent), y=percent)) + 
    geom_bar(stat="identity")+
  scale_fill_viridis(discrete=TRUE)+
    # Set up the plot theme
  theme_bw(base_size = 15)+
  theme(panel.grid = element_blank())+ 
  geom_hline(yintercept = 0)+
    labs(fill = "")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(x = "", y = "Percent Change")+
  theme(strip.text.x = element_blank(), 
          strip.background = element_blank()) 


ggsave("Marine Traffic - Territories.pdf", 
       dpi = 600, 
       width = 450, height = 200, unit = "mm")
    
```



## Modelling the overall percent change across all countries

The model for Apple data has been run in the Apple code.

```{r}

## Marine Traffic
boat <- read.csv("eez_ais_change.csv")

boat <- boat %>% filter(TERRITORY1 != "Heard and McDonald Islands")

marine.mod <- lme(data = boat, change~1, random=~1|TERRITORY1)
summary(marine.mod)

# Intercept: -9.393568
# Error: 1.881711
# df: 246
# t-val = -4.992035


## Air Traffic
air.mod <- lme(data = air, percent~1, random=~1|country)
summary(air.mod)


## Residential

#reverse residential percentage
googlepark$res.rev <- googlepark$residential*-1

mod.res <- gamm(data = googlepark, res.rev ~ lockdown + weekend + s(julian), random = list(country=~1))
# Intercept: -10.540777
# Error: 0.2502575
# df: 9157
# t-val = -42.11972


## Park
mod.park <- gamm(data = googlepark, parks ~ lockdown + weekend + s(julian), random = list(country=~1))

summary(mod.park$gam)

# Intercept: -19.6368
# Error: 0.9766
# df: 9193
# t-val = -20.106

```
